---
title: "HW2"
author: "This author section should be left blank empty for anonymous grading"
date: "Due 2024-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Here you load the necessary packages.  
library(tidyverse)


# Here you enter your name: Soju Hokari - It won't show in the .html file but show in the .rmd file. 
```

## Instruction: 
Once you have completed your answer, please submit both the knitted .html file and the .rmd file together.

### Question 1: Write your code in the code chunk and your text answer after the code chunk.

```{r}
fruitfly <- read.csv("datasets/fruitfly.csv")
summary(fruitfly)
```
The mean lifespan of all of the fruitflies is `r mean(fruitfly$lifespan)` days. The standard deviation of lifespan is `r sd(fruitfly$lifespan)` days.

```{r}
ggplot(fruitfly) +
  geom_histogram(binwidth = 5, aes(x = lifespan)) +
  ylab("Count") +
  xlab("Lifespan (days)")
```

The distribution of the overall lifespan looks to be relatively normal.


### Question 2: Write your code in the code chunk and your text answer after the code chunk.

```{r}
ggplot(fruitfly) +
  geom_point(aes(x = thorax, y = lifespan)) +
  ylab("Lifespan (days)") +
  xlab("Length of thorax (mm)")

cor <- cor(fruitfly$thorax, fruitfly$lifespan)
```
It looks like there is a linear relationship. The correlation coefficient is `r cor`, which indicates that there is a moderate positive relationship between these two variables.

### Question 3: Write your code in the code chunk and your text answer after the code chunk.

```{r}
y <- fruitfly$lifespan
x <- fruitfly$thorax
lm <- lm(y ~ x)
summary(lm)
```

The slope of the model is `r lm$coefficients[2]`. This slope implies that for every 1 additional mm in length of a fly's thorax, the lifespan of that fly will be `r lm$coefficients[2]` days longer.

### Question 4: Write your code in the code chunk and your text answer after the code chunk.

H0: slope ($\beta_1$) == 0
Ha: slope ($\beta_1$) != 0

```{r}
slope_standard_error <- summary(lm)$coefficients[2, 2]

t_statistic <- (lm$coefficients[2] - 0) / slope_standard_error

p_value <- 2 * pt(t_statistic, length(fruitfly$lifespan) - 2, lower.tail = F)
```

The p-value is `r p_value`, which is less than 0.05, so we can reject the null hypothesis, meaning that the slope is NOT zero, and the linear relationship between _lifespan_ and _thorax_ is **significant**.

### Question 5: Write your code in the code chunk and your text answer after the code chunk.

```{r}
# Using the formula:
critical_value <- qt(0.05, length(fruitfly$lifespan) - 2,lower.tail = F) # Find the critical value for 95% CI

upper <- lm$coefficients[2] + critical_value * slope_standard_error
lower <- lm$coefficients[2] - critical_value * slope_standard_error

# Using the R function `confint()`
confidence_interval <- confint(lm, level = 0.9)
```

Using the formula, the 90% confidence interval is [`r lower`, `r upper`].

Using the function, the 90% confidence interval is [`r confidence_interval[2, 1]`, `r confidence_interval[2, 2]`]

### Question 6: Write your code in the code chunk and your text answer after the code chunk.


```{r}
# Find individual fruitfly lifespan with thorax value 0.8 mm
individual <- predict(lm, newdata = data.frame(x = 0.8), se.fit = TRUE, interval = "prediction")

# Find average fruitfly lifespan with thorax value 0.8 mm
confidence <- predict(lm, newdata = data.frame(x = 0.8), se.fit = TRUE, interval = "confidence")
```

The predicted value of lifespan is `r individual$fit[1]` days.

The prediction interval for an individual fruitfly's lifespan with a 0.8mm thorax is [`r individual$fit[2]` days, `r individual$fit[3]` days].

The confidence interval for the average fruitfly's lifespan with a 0.8mm thorax is [`r confidence$fit[2]` days, `r confidence$fit[3]` days].

### Question 7: Write your code in the code chunk and your text answer after the code chunk.

```{r}
x_values <- data.frame(x = seq(0.6,1,0.01)) # seq() generates numbers from 10 to 90 by increments of 5

confidence_intervals <- predict(lm, newdata = x_values, interval = "confidence")
prediction_intervals <- predict(lm, newdata = x_values, interval = "prediction")

matplot(
  x_values$x,
  cbind(confidence_intervals, prediction_intervals[,-1]), # in order to avoid having two `fit` columns, we will extract the `fit` column from prediction_intervals
  lty = c(1,2,2,3,3),
  type = "l",
  col = c("black", "blue", "blue", "red", "red"),
  ylab = "predicted y"
) # lty = line type, type = line, ylab = y axis label
```
