---
title: "Lecture06Practice - MLR Inference"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```


## 1. Significance tests: Overall F test, partial F test, and individual t test.

Working on the dataset Prestige
```{r}
lmfull <- lm(prestige ~ income + education, data = Prestige)
summary(lmfull)

# Interpretation of the coefficient estimate: While education is held at a
#       constant level, q additional dollar increase in income is associated
#       with an increase in mean prestige score of 0.00136.
# 1000 increase in income ~ 0.136 increase in prestige score
# 1 year increase in education is associated with an increase in mean prestige
#       score of 4.13, while other variables are held at a constant level.
```

```{r}
# We're not going to use the ANOVA functions...
# IGNORE THE FOLLOWING.
anova(lmfull)
Anova(lmfull, type = "II")
Anova(lmfull, type = "III")
```

## Overall F-test

$\beta

```{r}
lmfull <- lm(prestige ~ income + education, data = Prestige)
lmnull <- lm(prestige ~ 1, data = Prestige) # Using one variable as a constant
#                                               to model the null hypothesis
summary(lmnull)

# Put the two models together in ANOVA
anova(lmnull, lmfull) # overall F test: comparing the full model and the null
#                                           model
```

Given the p-value, we can reject the null. So at least one of these predictors are significant predictors in the model.
Another way to perform overall F-test: Run summary of lmfull:

```{r}
summary(lmfull)
```

The bottom line has f-statistics, which tells us the same result as the ANOVA above.

## Partial F test

Full model: y = $\beta$~0~ + $\beta$~education~X~education~ + $\beta$~income~X~income~ + $\epsilon$

H0: $\beta$~income~ = 0 (income is a significant predictor)

So the reduced model is y = $\beta$~0~ + $\beta$~education~X~education~ + $\epsilon$

```{r}
# test coefficient of income
lmreduced <- lm(prestige ~ education, data = Prestige)


anova(lmreduced, lmfull) # partial F test: comparing the full model and the
#                                           reduced model

summary(lmfull) # individual t test: for one slope
```

F-statistic is 36.856 and P-value is 2.355e-8, so we reject the null. Income is a significant predictor variable.



