---
title: "Lecture02 practice"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(ggplot2)
#library(car)

```

## Import dataset to R

```{r}

Duncan<-read.table("http://socserv.socsci.mcmaster.ca/jfox/Books/Applied-Regression-3E/datasets/Duncan.txt",header=T)

#or 

Duncan<-read.table("Duncan.txt") 
# If you save/download the dataset and Rmarkdown file at the same folder, then you can simplify the above code to import the data

#or

library(car)
Duncan
```


## Headers start with hashtags

# Header 1

## Header 2

### Header 3

#### Header 4

## Making a table

| Col1 | Col2 | Col3 |
|------|------|------|
|      |      |      |
|      |      |      |
|      |      |      |

| .   | .   |
|-----|-----|
| .   | .   |

| Variable | Mean |
|---------:|:-----|
|    Girth | 13.1 |

## Lists

-   unordered list
    -   sub-item 1
    -   sub-item 2
        -   sub-sub-item 1
-   item 2

1.  ordered list
2.  item 2
    i)  sub-item 1 A. sub-sub-item 1

Check Rmarkdown cheat sheet!!!

## inline R code

```{r}
set.seed(123) # Guarantee the same result for random stuff

x <- sample(1:10,5, replace=F)
x
m<-mean(x)
```

inline code, using backquote + r

The value of `r m` is the average of one sample.

## Statistics

### Variables of Duncan:

-   type: Type of occupation: prof, professional and managerial; wc, white-collar; bc, blue collar.
-   income: Percent of males in occupation earning \$3500 or more in 1950.
-   education: Percent of males in occupation in 1950 who were high-school graduates.
-   prestige: Percent of raters in NORC study rating occupation as excellent or good in prestige.

### Extract observations from a dataset

```{r}
Duncan$income

# index operator to select certain columns and rows
Duncan[,2]

# Including a logical expression, to select only "prof" values
Duncan$income[Duncan$type == "prof"]  # including the logical expression

```

### Numerical Summaries

average and sd

```{r}
mean(Duncan$income[Duncan$type=="bc"])
sd(Duncan$income)

# Calculate the mean income per type
tapply(Duncan$income, Duncan$type, mean)

# Calculate the sd income per type
tapply(Duncan$income, Duncan$type, sd)
```

### Visualization

```{r}
# one variable
hist(Duncan$income)
boxplot(Duncan$income)

# two variables
#plot(x,y)
# plot(y~x)
plot(Duncan$income, Duncan$prestige)

# Aesthetics!
plot(Duncan$income, Duncan$prestige, main="The scatterplot", type="p", pch=4, cex=2, col="red")   
```

More complicated plots

```{r}
plot(Duncan$education,Duncan$income,type="n",xlab="education",ylab="income") ## Don't plot points.
points(Duncan$education[Duncan$type=="bc"],Duncan$income[Duncan$type=="bc"],pch=15, col=1)
points(Duncan$education[Duncan$type=="prof"],Duncan$income[Duncan$type=="prof"],pch=16,col=2)
points(Duncan$education[Duncan$type=="wc"],Duncan$income[Duncan$type=="wc"],pch=17,col=4)
legend(10,70,c("bc","prof","wc"),col=c(1,2,4),pch=c(15,16,17))
```

```{r}
library(ggplot2)
ggplot(Duncan, aes(x=education, y=income))+
  geom_point(aes(col=type))
```

### Inferential Statistics

Construct a confidence interval and do a hypothesis testing for the average income

```{r}
xbar<-mean(Duncan$income)
s<-sd(Duncan$income)
n<-nrow(Duncan)
n
alpha<-0.05
criticalt<-qt(0.025,n-1,lower.tail = F)
```

A confidence interval for true mean of income

```{r}
L<-xbar-criticalt*s/sqrt(n)
U<-xbar+criticalt*s/sqrt(n)
```

A 95% CI for mean is (`r L`, `r U`).

Hypothesis testing

```{r}
# method 1
t<-(xbar-50)*sqrt(n)/s
2*pt(t,n-1,lower.tail = T)

# method 2
t.test(Duncan$income, alternative = "two.sided", mu=50, conf.level = 0.95)

```

### Check Normality

```{r}
qqnorm(Duncan$income)
qqline(Duncan$income)

qqnorm(Duncan$prestige)
qqline(Duncan$prestige)

```

```{r}
plot(Duncan$income, Duncan$prestige)
cor(Duncan$income,Duncan$prestige)

```
