---
title: "HW1"
author: "This author section should be left blank empty for anonymous grading"
date: "Due 2024-01-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Here you load the necessary packages.
library("ggplot2")
library("tidyverse")

# Here you enter your name: Soju Hokari - It won't show in the .html file but show in the .rmd file. 
```

## Instruction: 
Once you have completed your answer, please submit both the knitted .html file and the .rmd file together.

### Question 1: Write your code in the code chunk and your text answer after the code chunk.

```{r}
y <- c(1,2,1,3,4,1,1,4,2,1,3,4,3,2,1,3,4,1,2,3,1,1,2,1,1,3,4)
y[y == 1] <- "Freshman"
y[y == 2] <- "Sophomore"
y[y == 3] <- "Junior"
y[y == 4] <- "Senior"
```

Variable `y` is now the following:

`r y`

### Question 2: Write your code in the code chunk and your text answer after the code chunk.

```{r}
y <- c(105,69,86,100,82,111,104,110,87,108,87,90,94,113,112,98,80,97,95,111,114,89,95,126,98)

xbar <- mean(y)

n <- length(y)

s <- sd(y)

# We're using a 90% conf. int., so we set alpha to 0.1
alpha <- 0.1

criticalt <- qt(alpha/2, n-1, lower.tail = F)

lower <- xbar - criticalt * s / sqrt(n)
upper <- xbar + criticalt * s / sqrt(n)

# OR Just use a t-test:
  # results <- t.test(y, alternative = "two.sided", mu = 100, conf.level = 0.9)

  # lower <- results$conf.int[1]
  # upper <- results$conf.int[2]
```

**The 90% confidence interval for the average student IQ in the school is [`r lower`, `r upper`]**

### Question 3: Write your code in the code chunk and your text answer after the code chunk.

```{r}
# H0: average IQ in the school is higher than 100
# Ha: average IQ in the school is not higher than 100

results <- t.test(y, alternative = "less", mu = 100, conf.level = 0.95)
```

The p-value is `r results$p.value`, which is more than 0.05. We therefore **fail
to reject the null hypothesis**, which means that the population mean for the IQ
of students at the school could be less than 100.

### Question 4: Write your code in the code chunk and your text answer after the code chunk.

```{r}
# Import `expenditure` data set; clean it up a little
expenditure <- read.table("datasets/expenditure.txt", header=T)

expenditure$Region <- factor(expenditure$Region, labels = c("Northeast", "North Central", "South", "West"))

summary(expenditure)

# plot it!
pairs(~ Y + X1 + X2 + X3, data = expenditure,
      labels = c("Per capita expenditure\non public education", "Per capita\npersonal income", "Residents per\nthousand under 18", "People per thousand\nin urban areas"),
      pch = ".",
      cex = 5,
      )
```
A scatterplot matrix is the appropriate graph to visualize these relationships. To me, it looks like there is a strong positive correlation between Y (per capita expenditure on public education) and X1 (per capita personal income), a weak positive correlation between Y and X3 (number of people per thousand residing in urban areas), and no correlation between Y and X2 (number of residents per thousand under 18 years of age). Between the X variables, there appears to be a positive correlation between X1 and X3, a weak negative relationship between X1 and X2, and what appears to be no relation between X2 and X3.

### Question 5: Write your code in the code chunk and your text answer after the code chunk.

```{r}
boxplot(expenditure$Y ~ expenditure$Region)

results <- t.test(expenditure$Y[expenditure$Region == "West"], alternative = "two.sided", mu = 100, conf.level = 0.95)
```

A side-by-side boxplot is most appropriate to visualize the relationship between Y and Region. On average, the West has the highest per capita expenditure on public education. The confidence interval for the true average per capita expenditure in the West is [`r results$conf.int`].

### Question 6: Write your code in the code chunk and your text answer after the code chunk.

```{r}
ggplot(expenditure) +
  geom_point(aes(x = X1, y = Y, color = Region, shape = Region)) +
  xlab("Per capita personal income") +
  ylab("Per capita expenditure on public education") +
  labs(color = "Region")
```

Y and X1 seem to have a relatively strong, positive, linear relationship.

