---
title: "LecturePractice05 - MLR Estimation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```


```{r}
??Prestige
lm1<-lm(prestige~income, data=Prestige)
summary(lm1)
```
income is a significant indicator of the prestige


```{r}
lm2<-lm(prestige~income+education, data=Prestige)
summary(lm2)
```

income and education are both significant predictors of prestige

# Partial regression plot to check education effect
```{r}
lm3 <- lm(education~income, data = Prestige)

plot(lm1$residuals~lm3$residuals, ylab = "residuals of regressing income on prestige\n(part of prestige that is not accounted for by income)", xlab = "residuals of regressing education on income\n(part of income that is not accounted for by education)")

# Run a linear model to regress lm1$residuals and lm3$residals
summary(lm(lm1$residuals~lm3$residuals))
```
The p-value is small enough to reject the null (the null is that there is no linear relationship between `lm1$residuals` and `lm3$residuals`). So we should include the education variable.

# plot with giving a predictor value in the lecture 5
```{r}
clock <- read.table("http://www.statsci.org/data/general/auction.txt", header = T)

# We can use fitted lines to show the relationship between two linear variables in 2d (instead of a plane)
# price^ = B0^ + B1^ * Bidder + B2^ * Age (fitted model)
# The above is a fitted plane. But if we fix one variable at a constant level, then we can represent it in 2D!

lm.out1 <- lm(Price~Bidders+Age, data = clock)
lm.out1$coefficients

plot(Price~Bidders, data = clock)
clock$Age # We will choose age = 120, age = 150, age = 170

plot(Price~Bidders, data = clock)
abline(lm.out1$coefficients[1] + 120 * lm.out1$coefficients[3], lm.out1$coefficients[2], col = 1)
abline(lm.out1$coefficients[1] + 150 * lm.out1$coefficients[3], lm.out1$coefficients[2], col = 2)
abline(lm.out1$coefficients[1] + 170 * lm.out1$coefficients[3], lm.out1$coefficients[2], col = 4)
legend(x=12,y=2000, c("Age=120","Age=150","Age=170"),col=c(1,2,4),pch=c(15,16,17))

plot(Price~Bidders, data = clock)
for (i in c(120, 130, 140, 150, 160, 170)) {
  abline(lm.out1$coefficients[1] + i * lm.out1$coefficients[3], lm.out1$coefficients[2], col = 1)
}
legend(x = 12, y = 2000, c("Age=120", "Age=130", "Age=140", "Age=150", "Age=160", "Age=170"), col=c(1, 2, 4, 5, 6, 7), pch=c(15, 16, 17, 18, 19, 20))


```

So B0^ = `r lm.out1$coefficients[1]`, B1^ = `r lm.out1$coefficients[2]`, and B2^ = `r lm.out1$coefficients[3]`.
> price^ = B0^ + B1^ * Bidder + B2^ * Age


